Test Suite Parallelization in Open-Source Projects:\\ A Study on Its Usage and Impact

### 2019-05-22 tse19/main.tex
pass on section 2

\Mar{We need to reorganize this a bit as to accomodate the new
  addition to the paper...}

@@

% SECTION 2
\section{\label{sec:background}Parallel Execution of Test Suites}
Figure~\ref{fig:levels} illustrates different levels where
parallelism in test execution can be obtained.
The highest level indicates
parallelism obtained through different machines on the
network.  For instance, using virtual machines from a cloud service to
\begin{wrapfigure}{r}{0.13\textwidth}  
  \vspace{-2mm}
  \centering
  %  \includegraphics[width=0.35\textwidth]{figs/parallel-levels.pdf}    
  \includegraphics[width=2cm]{figs/parallel-levels-short.pdf}  
  \caption{\label{fig:levels}Levels of parallelism.}
  \vspace{-2mm}
\end{wrapfigure}

distribute test execution.  The lowest levels denote parallelism
obtained within a single machine.  These levels are complementary:
the lowest levels leverage the computing power of server
nodes whereas the highest level leverages the aggregate processing
power of a network of machines.
This paper focuses on low-level parallelism, where computation can be
offloaded at different CPUs within a machine and at different threads
within each CPU.  This form of parallelism is enabled through build
systems (spawning processes in different CPUs) and testing frameworks
(spawning threads in one given CPU).  It is important to note that a variety
of testing frameworks provide today support for parallel test
execution (e.g., JUnit~\cite{junit-org}, TestNG~\cite{testng}, and
NUnit~\cite{nunit}) as to benefit from the available power of popular multi-core processors.
In the following, we elaborate relevant features of testing frameworks
and build systems for parallelization.  We focused on Java, Maven, and JUnit but the
discussion can be generalized to other language and tools.

@@

\begin{itemize}
\item
    \textbf{Sequential (\Seq).}~No parallelism is involved.
\item
    \textbf{Sequential classes; parallel methods
      (\SeqClassParMeth).}~This configuration corresponds to running
    test classes sequentially, but running test methods from those
    classes concurrently.
\item
    \textbf{Parallel classes; sequential methods
      (\ParClassSeqMeth{}).}~This configuration corresponds to running
    test classes concurrently, but running test methods sequentially.
\item
    \textbf{Parallel classes; Parallel methods
      (\ParClassParMeth).}~This configuration runs test classes and
    methods concurrently.\Comment{corresponds to the
    combination of \SeqClassParMeth{} and \ParClassSeqMeth{}.  }
\end{itemize}

Notice that an important aspect in deciding which configuration to use (or in
designing new test suites) is the possibility of race conditions on shared data
during execution.  Data sharing can occur, for example, through state that is
reachable from statically-declared variables in the program or through variables
declared within the scope of the test class or even through resources available
on the file system and the network~\cite{luo-etal-fse2014}.  Considering data
race avoidance, configuration \SeqClassParMeth{} is preferable over
\ParClassSeqMeth{} when it is clear that test methods in a class do not
manipulate shared state, which can be challenging to
determine~\cite{bell-etal-esecfse2015}.  Similarly, \ParClassSeqMeth{} is
preferable over \SeqClassParMeth{} when it is clear that several test methods in
a class perform operations involving shared data.  Configuration
\ParClassParMeth{} does not restrict scheduling orderings.  Consequently, it is
more likely to manifest data races during execution. Note that speedups depend
on several factors, including the test suite size and distribution of test
methods per class.

@@

\begin{itemize}
\item
  \textbf{Forked JVMs with sequential methods (\ForkSeq).}~The build
  system spawns multiple JVMs with this configuration, assigning a
  partition of the set of test classes to each JVM.  Test classes and methods
  run sequentially within each JVM.
\item
  \textbf{Forked JVMs with parallel methods (\ForkParMeth).}~With
  this configuration, the build system forks multiple JVMs, as
  \ForkSeq{} does, but it runs test methods concurrently, as
  \SeqClassParMeth{} does.
\end{itemize}

@@

Note from the listing that forking can only be combined with
configuration \SeqClassParMeth{} (see Section~\ref{sec:frameworks}) as
Maven made the design choice to only accept one test class at a time
per forked process.  Maven offers an option to reuse JVMs that can be
used to attenuate the potentially high cost of spawning new JVM
processes on every test class (if reuse is enabled) and also to
achieve test isolation (if reuse is disabled).

%% \Mar{$\leftarrow$Are you sure about this?
%%   It seems silly.  Maybe, the rationale is to ``keep design
%%   simple''.}\Jbc{I'm sure. in addition to keep design simple, I feel
%%   like this was initially conceived to run tests in isolation}
%% \Mar{$\leftarrow$confirma?  se sim, pode mostrar isto na
%%   configuracao abaixo?}\Jbc{added...}
%%We are unaware of other build system capable of running multiple
%%classes at time within a forked process.
%%  to
%% define tasks related to the project building and these tasks are
%% performed several plugins.This file
%% contains defintions of tasks, which are implemented by a collection of
%% plugins.
%% \Jbc{falar do surefire e explicar a
%% configuracao ilustrada - Surefire levanta uma JVM por core e cria uma
%% pool de 5 threads em cada JVM para executar testes em paralelo.}
\begin{figure}[h!]
\centering
\scriptsize
\lstset{
  escapeinside={@}{@},
  numbers=none,xleftmargin=1em,frame=none,framexleftmargin=0.5em,
  basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
  morekeywords={parallel, threadCount, perCoreThreadCount,
  forkCount, reuseForks},
  deletekeywords={true}
}
\vspace{-4ex}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <forkCount>1C</forkCount>
        <reuseForks>true</reuseForks>
        <parallel>methods</parallel>
        <threadCount>5</threadCount>
    </configuration>
</plugin>
\end{lstlisting}
  \vspace{-4ex}
  \caption{\label{fig:surefire} Configuration \ForkParMeth{} on
    Maven.}
\end{figure}

\subsubsection*{Example}~Figure~\ref{fig:surefire} shows a
fragment of a Maven configuration file, known as \pomf{}, highlighting
options to run tests using the parallel execution mode \ForkParMeth{}.
Maven implements this feature through its Surefire JUnit test
plugin~\cite{maven-surefire-plugin}.  With this configuration, Maven
forks one JVM per core (\CodeIn{forkCount} parameter) and uses five
threads (\CodeIn{threadCount} parameter) to run test methods
(\CodeIn{parallel} parameter) within each forked JVM.  Maven reuses
created JVMs on subsequent forks when execution of a test class
terminates (\CodeIn{reuseFork} parameter).

>>>>

\Mar{We need to reorganize this a bit as to accomodate the new
  addition to the paper...Revisit this later...}

@@

% SECTION 2
\section{\label{sec:background}Parallel Execution of Test Suites}

\begin{wrapfigure}{r}{0.13\textwidth}  
  \vspace{-2mm}
  \centering
  %  \includegraphics[width=0.35\textwidth]{figs/parallel-levels.pdf}    
  \includegraphics[width=2cm]{figs/parallel-levels-short.pdf}  
  \caption{\label{fig:levels}Levels of parallelism.}
  \vspace{-2mm}
\end{wrapfigure}
This section explains basic concepts related to test parallelization
in Java projects. Figure~\ref{fig:levels} illustrates different levels
of parallelism in test execution.  The highest level indicates
parallelism obtained through different machines on the network. For
instance, using virtual machines from a cloud service to distribute
test workload. The Google TAP~\cite{google-tap,google-ci} and the
Microsoft CloudBuild~\cite{prasad-shulte-ieee-microsoft-ci} systems
are examples of this kind of parallelization. The lowest levels
illustrate parallelism obtained within a single machine. These levels
are complementary to each other: the lowest levels leverage the
computing power of individual nodes whereas the highest level
leverages the aggregate processing power of a network of
machines. This paper focuses on low-level parallelism, where
computation can be offloaded at different CPUs within a machine and at
different threads within each CPU.  This form of parallelism is
enabled through build systems---by spawning processes in different
CPUs---and testing frameworks---by spawning threads in one given CPU.
It is important to note that a variety of testing frameworks provide
today support for parallel test execution (e.g.,
JUnit~\cite{junit-org}, TestNG~\cite{testng}, and NUnit~\cite{nunit})
as to benefit from the available power of popular multi-core
processors.  In the following, we elaborate relevant features of
testing frameworks and build systems for parallelization.  We focused
on Java, Maven, and JUnit but the discussion can be generalized to
other language and tools.

@@

\begin{itemize}
\item
    \textbf{Sequential (\Seq).}~No parallelism is involved.
\item
    \textbf{Sequential classes; parallel methods
      (\SeqClassParMeth).}~This configuration corresponds to running
    test classes sequentially, but running the test methods from those
    classes concurrently.
\item
    \textbf{Parallel classes; sequential methods
      (\ParClassSeqMeth{}).}~This configuration corresponds to running
    test classes concurrently, but running test methods sequentially.
\item
    \textbf{Parallel classes; Parallel methods
      (\ParClassParMeth).}~This configuration runs test classes and
    methods concurrently.\Comment{corresponds to the
    combination of \SeqClassParMeth{} and \ParClassSeqMeth{}.  }
\end{itemize}

An important aspect to consider when designing new test suites or when
deciding which configuration to use for existing test suites is the
possibility of race conditions on shared data during test
execution. Although conceptually test cases are independent, they
often share data\Fix{cite Mike's paper, bell-etal-esecfse2015}. Data
sharing can occur, for example, through state that is reachable from
statically-declared variables in the program or through variables
declared within the scope of the test class or even through resources
available on the file system and the network~\cite{luo-etal-fse2014}.

Test flakiness and execution speedups depend on a number of
factors. Considering flakiness, a configuration may be preferrable to
another depending on the characteristics of the project and its test
suite. For example, configuration \ParClassSeqMeth{} may be preferable
over \SeqClassParMeth{} when it is clear that several test methods in
a given class perform operations involving shared data. As
configuration \ParClassParMeth{} does not restrict scheduling
orderings, it is more likely to manifest data races during execution
if there are sources of data sharing across test classes and their
methods. Note also that speedups depend on several factors, including
the test suite size and the distribution of test methods per class.

@@

\begin{itemize}
\item
  \textbf{Forked JVMs with sequential methods (\ForkSeq).}~The build
  system spawns multiple JVMs with this configuration, assigning a
  partition of the set of test classes to each JVM.  Test classes and methods
  run sequentially within each JVM as \Seq\ does.
\item
  \textbf{Forked JVMs with parallel methods (\ForkParMeth).}~As
  \ForkSeq{}, the build system forks multiple JVMs with this
  configuration, but it runs test methods concurrently, as
  \SeqClassParMeth{}.
\end{itemize}

@@

Note from the listing that forking can only be combined with
configuration \SeqClassParMeth{} (see Section~\ref{sec:frameworks}) as
Maven made the design choice to only accept one test class at a time
per forked process. Intuitively, this option enables safe execution of
tests when dependent methods exist across different classes\Fix{cite
  Mike's work}. For the cases the developer prepared code against that
threat, Maven offers an option to ``reuse'' JVMs as to avoid the
non-negligible cost of creating JVM instances.

%% \Mar{$\leftarrow$Are you sure about this?
%%   It seems silly.  Maybe, the rationale is to ``keep design
%%   simple''.}\Jbc{I'm sure. in addition to keep design simple, I feel
%%   like this was initially conceived to run tests in isolation}
%% \Mar{$\leftarrow$confirma?  se sim, pode mostrar isto na
%%   configuracao abaixo?}\Jbc{added...}
%%We are unaware of other build system capable of running multiple
%%classes at time within a forked process.
%%  to
%% define tasks related to the project building and these tasks are
%% performed several plugins.This file
%% contains defintions of tasks, which are implemented by a collection of
%% plugins.
%% \Jbc{falar do surefire e explicar a
%% configuracao ilustrada - Surefire levanta uma JVM por core e cria uma
%% pool de 5 threads em cada JVM para executar testes em paralelo.}
\begin{figure}[h!]
\centering
\scriptsize
\lstset{
  escapeinside={@}{@},
  numbers=none,xleftmargin=1em,frame=none,framexleftmargin=0.5em,
  basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
  morekeywords={parallel, threadCount, perCoreThreadCount,
  forkCount, reuseForks},
  deletekeywords={true}
}
\vspace{-4ex}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <forkCount>1C</forkCount>
        <threadCount>5</threadCount>    
        <parallel>methods</parallel>
        <reuseForks>true</reuseForks>        
    </configuration>
</plugin>
\end{lstlisting}
  \vspace{-4ex}
  \caption{\label{fig:surefire} Fragment of a Maven pom.xml file with
    an \ForkParMeth{} configuration.}
\end{figure}

\subsubsection*{Example}~Figure~\ref{fig:surefire} shows a
fragment of a Maven configuration file, known as \pomf{}, highlighting
options to run tests using the parallel execution mode \ForkParMeth{}.
Maven implements this feature through its Surefire JUnit test
plugin~\cite{maven-surefire-plugin}.  With this configuration, Maven
forks one JVM per core (\CodeIn{forkCount} parameter) and uses five
threads (\CodeIn{threadCount} parameter) to run test methods
(\CodeIn{parallel} parameter) within each forked JVM. In addition, Maven reuses
created JVMs on subsequent forks when execution of a test class
terminates (\CodeIn{reuseFork} parameter). This is an example of a
very aggressive setup, \ie{}, a setup that could substantially speedup
execution but could introduce test flakiness if test code is not
prepared for parallel execution.
